Parsing Postfix log files

### 2009-03-27 doc/logparser-implementation.tex
Wrote introduction for implementation chapter.

XXX WRITE AN INTRODUCTION\@.  MOST OF THIS IS CRAP, BUT I MIGHT BE ABLE TO
RESCUE SOME OF IT\@.  IT SHOULD HAVE A BLURB, FOLLOWED BY A PREVIEW OF THE
CHAPTER\@.

The parser deals with all the complications of parsing, the eccentricities
and oddities of Postfix log files, presenting the resulting data in a
normalised, simple to use representation.  The parser's task is to follow
the journey each mail takes through Postfix, combining the data captured by
rules into a coherent whole, saving it in a useful and consistent form, and
performing housekeeping duties.

There is a similarity between this design and the event-driven programming
paradigm commonly used in GUI programs, where one part of the program
responds to events (mouse clicks in a GUI program, log lines being
recognised in \parsername{}) and invokes the correct action.

@@

This section has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the resemblance between a database schema and an
\acronym{API} is described, followed by a detailed description and a
diagram of the database schema.  The framework is documented next,
including the steps it takes during initialisation, the parsing process,
and the conveniences it offers to users.  The data collected by the
framework is described, as are the ways in which various optimisations can
be disabled to demonstrate their effect, and the debugging options the
framework provides.  The implementation of actions in \parsername{} is
documented, including how frequently each action is specified by rules, and
brief descriptions of why some actions are more popular than others.  All
the actions that are part of \parsername{} are described in detail,
followed by the process of adding a new action to the parser.  The final
component of the architecture, the rules, is also the most visible
component, and its implementation is examined in detail with: an example
rule and an explanation of every field; a description of how to add new
rules and determine the values that should be used for each field; an
explanation of the algorithm used by the utility that creates new regexes
based on unparsed log lines, and how it differs from the original algorithm
it is based on.  How \parsername{} uses rule conditions and overlapping
rules is discussed, accompanied by a description of the regex snippets the
framework provides to simplify the regexes used in rules.  The many
complications and difficulties encountered while writing \parsername{}, and
the solutions developed to overcome them, are documented in depth, also
documented are how some solutions interact and which action or actions the
solution is implemented in.  This chapter concludes with a list of the
limitations identified in \parsername{}, followed by some possible
improvements suggested by the author.

>>>>

This chapter documents the implementation of \parsername{}, based on the
architecture described in \sectionref{parser architecture}.  Any design may
look good when examined in the abstract, but the real test of the design
comes with the first concrete implementation; only then do any practical
difficulties come to light.  Implementing this architecture was relatively
straightforward, with only minor changes required: rule conditions and
cascaded parsing were added, although later the implementation stopped
using cascaded parsing.  The difficulties came not from the architecture
and design, but from anomalies in the log files and Postfix's behaviour, as
described in \sectionref{complications}.

\parsername{} deals with all the complications of parsing, the
eccentricities and oddities of Postfix log files, presenting the resulting
data in a normalised, simple to use representation.  The parser's task is
to follow the journey each mail takes through Postfix, combining the data
captured by rules into a coherent whole, saving it in a useful and
consistent form, and performing housekeeping duties.

This chapter begins with the assumptions under which \parsername{} was
designed and written, followed by a flowchart showing the most common paths
taken through Postfix and \parsername{}, with a description of the stages
and transitions.  The topic of the next section is the database that
provides storage for the parser: any future work that analyses gathered
data will do so using the database, so the database schema acts as an
\acronym{API}.  A diagram of the database schema is provided, plus
documentation for every table and field.

The next three sections document the implementation of the three components
of the architecture.  First is the framework, including its initialisation
phase, the parsing process, and the conveniences it offers to users; that
section finishes with the features that are important for this thesis but
not strictly necessary for parsing: the performance data collected by the
framework, the optimisations that can be disabled to show their effect, and
the debugging options it provides.  The second component of the
architecture is the actions: their implementation is documented, with a
graph of how often each action is specified by rules, and a description of
why some actions are more popular than others.  Every action that is part
of \parsername{} is documented, and the actions section concludes with the
process of adding a new action, and how to determine the values for each of
the required fields.  Rules are the final component of the architecture,
and also the most visible to users, because it is likely that users will
need to add their own rules to recognise their log lines.  An example rule
used by \parsername{} is shown, with every field clearly documented; this
is followed by a description of adding new rules, and how to determine the
value of each of the fields.  \parsername{} provides a utility to create
regexes from unrecognised log lines; the algorithm it uses is documented,
including the differences between it and the original algorithm.  The rules
section finishes with a discussion of how \parsername{} uses rule
conditions and overlapping rules, plus a description of the regex snippets
provided to ease the process of writing regexes.

On first inspection, Postfix log files look like they will be simple to
parse, but this is not the case.  The many complications and difficulties
encountered while writing \parsername{}, and the solutions developed to
overcome them, are documented in \sectionref{complications}.  This chapter
finishes with a list of \parsernames{} limitations, and some possible
improvements that could be implemented.

@@

This chapter has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the resemblance between a database schema and an
\acronym{API} is described, followed by a detailed description and a
diagram of the database schema.  The framework is documented next,
including the steps it takes during initialisation, the parsing process,
and the conveniences it offers to users.  The performance data collected by
the framework is described, as are the ways in which various optimisations
can be disabled to demonstrate their effect, and the debugging options the
framework provides.  The implementation of actions in \parsername{} is
documented, including how frequently each action is specified by rules, and
brief descriptions of why some actions are more popular than others.  All
the actions that are part of \parsername{} are described in detail,
followed by the process of adding a new action to the parser.  The final
component of the architecture, the rules, is also the most visible
component, and its implementation is examined in detail with: an example
rule and an explanation of every field; a description of how to add new
rules and determine the values that should be used for each field; an
explanation of the algorithm used by the utility that creates new regexes
based on unparsed log lines, and how it differs from the original algorithm
it is based on.  How \parsername{} uses rule conditions and overlapping
rules is discussed, accompanied by a description of the regex snippets the
framework provides to simplify the regexes used in rules.  The many
complications and difficulties encountered while writing \parsername{}, and
the solutions developed to overcome them, are documented in depth; also
documented are how some solutions interact, and which action or actions the
solution is implemented in.  This chapter concludes with a list of the
limitations identified in \parsername{}, followed by some possible
improvements suggested by the author.
